Impact of asymmetric tail-dependence in extreme climatic events on extinction risk in population ecology

### 2019-03-19 MainText.Rmd
keywords added
plot changed notation sigma for standard deviation

\textbf{\textit{Keywords:}}

@@

To test our hypothesis from (Q1), we repeated the above analysis 
considering the vital rates coming from a 
an asymmetric copula with specific 
left (Clayton, Survival Gumbel)
or right (Survival Clayton, Gumbel) tail dependence
or from an extreme tail-depentdent one, 
when vital rates were considered to be affected
by extreme climatic events. A $N \times 2$ matrix, M, was generated
using `BiCopSim` function
from `VineCopula` package 
for each of the left 
(Clayton, Survival Gumbel) and right 
(Survival Clayton, Gumbel) tail-dependent bivariate copulas, 
where $N = numsteps(=25) \times numsims(=10000)$.
We kept the same Kendall's correlation ($\tau = 0.6$) for 
each bivariate copula from one parameter family which used its parameter accordingly
(see `iTau` function from `copula` package). Entries within each clumn of
matrix $M$, were thus 
independent of each others but 
the two columns of $M$ had a specified Kendall's 
correlation ($\tau = 0.6$) and underlying tail-dependence. 
Similarly, for extreme-tail dependence
(either left or right tail) a matrix M with two column was generated as described in 
Appendix \ref{SM-extr_tail_dep}.

@@

The environmental noises from any two patch, say, 
$(\epsilon_t^i, \epsilon_t^j)$ were normally 
distributed within each patch with mean $m=0$ and standard deviation $sd$, 
(see the specified value $sd$ for each model in 3$^{rd}$ 
column of Table \ref{tab_model_summary}), were 
independent through time, and showed the same 
spatial correlations between patches (to within sampling error) for every simulation. 
But they were made to exhibit extreme right- or extreme left-tail dependence between patches in
different simulations. A three dimensional array, numsims $\times$ n $\times$ numsteps, 
$\epsilon$, was made from these extreme tail-dependence noise 
matrix generated via algorithm stated in 
Appendix \ref{SM-extr_tail_dep} and used to calculate extinction risk 
for each of the specified model in Table \ref{tab_model_summary} for $numsteps = 25$ years, $n =$ 5 patches
and $numsims = 10000$ simulations.

@@

<!--test hypothesis from Q2 : ext risk with density-dependence metapopulation model with r from different regime-->
To test our hypothesis from (Q2) of \nameref{intro} for each metapopulation model,
we estimated extinction risk for a wide range of parameter $r$ 
which covered the undercompensatory and overcompensatory regime. We 
chose equispaced values of $r$ from the 95\% quantiles 
of possible range ($r_{min},r_{bf}$), where $r_{min}$ was the minimum value of $r$ 
for which the model equilibrium $P_e$ had always a finite, positive initial 
population (for details, see Appendix \ref{SM-model_stability}). For each chosen $r$ from each 
model where metapopulation could disperse locally with varying dispersal 
rate $d= 0,0.1,0.2,..0.5$, we calculated extinction risk with left 
tail dependent and right tail dependent environmental noise 
in different simulations, respectively. We also repeated the analysis 
for global dispersal for each metapopulation model. Script 
[`mydispersal.R`](https://github.com/sghosh89/ERC/blob/master/mydispersal.R) was used for this 
analysis.

@@

<!--Summary tables for 6 nonlinear models used-->
```{r read_tab_model_summary,echo=F, results='asis',message=F}
library(kableExtra)
library(dplyr)
dt<-readRDS("./Results/tab_model_summary.RDS")
knitr::kable(dt, escape=F,
             format="latex",digits=3, align="c",table.env='table*',
             caption = "Summary of six density dependence metapopulation models considered; last three columns indicated equilibrium population value ($P_e^{i}$), critical growth rate ($r_c^{i}$) where deterministic analog of models showed transition from under- to over-compensatory dynamics and bifurcation parameter ($r_{bf}^{i}$) where $P_e^{i}$ became unstable from stable steady state, respectively. Superscript $i=1,2,...,n$ indicates states for $i^{th}$ patch.\\label{tab_model_summary}",
             booktabs=T, linesep = "\\addlinespace"
             )%>%column_spec(c(1,3,5,6),width="1.5 cm")%>%kable_styling(latex_options = "basic",position = "center")%>%column_spec(2, width = "3.8 cm")
```

>>>>

\textbf{\textit{Keywords:}} extinction risk, extreme events, tail dependence, copula, density-dependence model

@@

To test our hypothesis from (Q1), we repeated the above analysis 
considering the vital rates coming from a 
an asymmetric copula with specific 
left (Clayton, Survival Gumbel)
or right (Survival Clayton, Gumbel) tail dependence
or from an extreme tail-depentdent one, 
when vital rates were considered to be affected
by extreme climatic events. A $N \times 2$ matrix, M, was generated
using `BiCopSim` function
from `VineCopula` package 
for each of the left 
(Clayton, Survival Gumbel) and right 
(Survival Clayton, Gumbel) tail-dependent bivariate copulas, 
where $N = numsteps(=25) \times numsims(=10000)$.
We kept the same Kendall's correlation ($\tau = 0.6$) for 
each bivariate copula from one parameter family which used its parameter accordingly
(see `iTau` function from `copula` package). Entries within each clumn of
matrix $M$, were thus 
independent of each others but 
the two columns of $M$ had a specified Kendall's 
correlation ($\tau = 0.6$) and underlying tail-dependence. 
Similarly, for extreme-tail dependent bivariant copula
(either left or right tail) a matrix M was generated from standard iid as described in 
Appendix \ref{SM-extr_tail_dep}.

@@

The environmental noises from any two out of $n$ patches, say, 
$(\epsilon_t^i, \epsilon_t^j)$ were normally 
distributed within each patch with mean $m=0$ and standard deviation $\sigma$ 
(see the specified value $\sigma$ for each model in 3$^{rd}$ 
column of Table \ref{tab_model_summary}). Noises were 
independent through time, and showed the same 
spatial correlations between patches (to within sampling error) for every simulation, 
but they were made to exhibit extreme right- or extreme left-tail dependence between patches in
different simulations. A three dimensional array, numsims $\times$ n $\times$ numsteps, 
$\epsilon$, was made from these extreme tail-dependence noise 
matrix generated via algorithm stated in 
Appendix \ref{SM-extr_tail_dep} and used to calculate extinction risk 
for each of the specified model in Table \ref{tab_model_summary} for $numsteps = 25$ years, $n =$ 5 patches
and $numsims = 10000$ simulations.

@@

<!--test hypothesis from Q2 : ext risk with density-dependence metapopulation model with r from different regime-->
To test our hypothesis from (Q2) of \nameref{intro} for each metapopulation model,
we estimated extinction risk for a wide range of parameter $r$, 
which covered the undercompensatory and overcompensatory regime. We 
chose equispaced values of $r$ from the 95\% quantiles 
of possible range ($r_{min},r_{bf}$), where $r_{min}$ was the minimum value of $r$ 
for which the model equilibrium $P_e$ had always a finite, positive initial 
population (for details, see Appendix \ref{SM-model_stability}). For each chosen $r$ from each 
model where metapopulation could disperse locally with varying dispersal 
rate $d= 0,0.1,0.2,..0.5$, we calculated extinction risk with left 
tail dependent and right tail dependent environmental noise 
in different simulations, respectively. We also repeated the analysis 
for global dispersal for each metapopulation model. Script 
[`mydispersal.R`](https://github.com/sghosh89/ERC/blob/master/mydispersal.R) was used for this 
analysis.

@@

<!--Summary tables for 6 nonlinear models used-->
```{r read_tab_model_summary,echo=F, results='asis',message=F}
library(kableExtra)
library(dplyr)
dt<-readRDS("./Results/tab_model_summary.RDS")
knitr::kable(dt, escape=F,
             format="latex",digits=3, align="c",table.env='table*',
             caption = "Summary of six density dependence metapopulation models considered at each patch; last three columns indicated equilibrium population value ($P_e^{i}$), critical growth rate ($r_c^{i}$ : transition point from under- to over-compensatory dynamics of deterministic models) and bifurcation parameter ($r_{bf}^{i}$ : transition point from stable to unstable equilibrium of deterministic models), respectively. Superscript $i=1,2,...,n$ indicates states for $i^{th}$ patch.\\label{tab_model_summary}",
             booktabs=T, linesep = "\\addlinespace"
             )%>%column_spec(c(1,3,5,6),width="1.4 cm")%>%kable_styling(latex_options = "basic",position = "center")%>%column_spec(2, width = "3.8 cm")
```
